A Classical Density-Functional Theory for Describing Water Interfaces

### 2011-03-04 paper/paper.tex
avoid redundant include
Tested-on: bingley (Linux 2.6.32-5-amd64)
make temperature a field rather than a scalar variable
Tested-on: bingley (Linux 2.6.32-5-amd64)

A large fraction of interesting chemistry---including all of molecular
biology---takes place in aqueous solution.  However, while quantum chemistry
now enables us to calculate the ground state energies of large molecules in
vacuum, prediction of the free energy of even the smallest molecules in the
presence of a solvent poses a continuing challenge due to the complex
structure of a liquid and the computational cost of \emph{ab initio}
molecular dynamics~\cite{car1985, grossman2004}.  The current state-of-the
art in \emph{ab initio} molecular dynamics is limited to around 64 water
molecules per unit cell, which places the minimum solute concentration at
0.84 molar and allows contact between the second solvation shell of
neighboring solutes~\cite{izvekov2005, choe2007}.  A more efficient
approach is needed in order to study nanoscale and hydrophobic solutes.

@@

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figs/finding-vapor-pressure}
\end{center}
\caption{By plotting the free energy vs. density it is possible to
  determine a saturated liquid or vapor state by the common tangent
  method.  }
\label{fig:homogeneous}
\end{figure}

>>>>

A large fraction of interesting chemistry---including all of molecular
biology---takes place in aqueous solution.  However, while quantum
chemistry now enables us to calculate the ground state energies of
large molecules in vacuum, prediction of the free energy of even the
smallest molecules in the presence of a solvent poses a continuing
challenge due to the complex structure of a liquid and the
computational cost of \emph{ab initio} molecular
dynamics~\cite{car1985, grossman2004}.  The current state-of-the art
in \emph{ab initio} molecular dynamics is limited to around 64 water
molecules per unit cell, which places the minimum solute concentration
at 0.84 molar and allows contact between the second solvation shell of
neighboring solutes~\cite{izvekov2005, choe2007}.  On top of this,
\emph{ab initio} approaches using classical molecular dynamics have
\textcolor{red}{vague problems}\cite{weber2010ab-initio-water}.  A
more efficient approach is needed in order to study nanoscale and
hydrophobic solutes.

@@

\subsection{Crossover formalism for handling critical behavior}

I am currently working on implementing the crossover approach of
Kiselev and Ely which is designed to make the critical behavior come
out right\cite{kiselev2006new}.  This method works by defining a
\emph{classical} free energy $f_0$---which where classical really is
closer to \emph{mean field} in meaning, and indicates a theory that
doesn't include long-range fluctuations.  This free energy is then
used to construct a corrected free energy by feeding portions of it a
``fake'' density and temperature that are constructed so as to
reconstruct the correct universal scaling laws near the critical
point.

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figs/finding-vapor-pressure}\\
\includegraphics[width=\columnwidth]{figs/near-critical-point}
\end{center}
\caption{By plotting the free energy vs. density it is possible to
  determine a saturated liquid or vapor state by the common tangent
  method.  }
\label{fig:homogeneous}
\end{figure}
